---
title: "Data Management Project"
author : Lou Dhiver and Lydie Lazure 
format: html
---
# Phase 2 of the project

### Github project :

https://github.com/loudhvr/DM-14-Dhiver-Lazure.git

### Sources for data sets :

-   World Health Organization (WHO): Ambient air pollution attributable death rate (per 100,000 population, age-standardized) : https://www.who.int/data/gho/data/indicators/indicator-details/GHO/ambient-air-pollution-attributable-death-rate-(per-100-000-population-age-standardized)
-   World Health Organization (WHO): Ambient air pollution attributable DALYs (per 100,000 population, age-standardized) : https://www.who.int/data/gho/data/indicators/indicator-details/GHO/ambient-air-pollution-attributable-dalys-(per-100-000-population-age-standardized)
-   World Bank : GDP per capita (current US\$): https://data.worldbank.org/indicator/NY.GDP.PCAP.CD
-   World Health Organization (WHO): Ambient air pollution, Concentrations of fine particulate matter (PM2.5) per country : https://apps.who.int/gho/data/view.main.SDGPM25116v

#### World Health Organization (WHO) :

The World Health Organization (WHO) is a global leader in the production and dissemination of reliable, standardized health-related data. As the health arm of the United Nations, WHO collaborates with 194 member states to collect, validate, and analyze health statistics, ensuring consistency and accuracy across diverse regions and health systems. Through its Global Health Observatory (GHO), WHO offers an extensive repository of data on key health indicators, including life expectancy, maternal and child health, disease prevalence, vaccination coverage, and environmental factors like air and water quality. These datasets are critical for monitoring global health trends, assessing progress toward the Sustainable Development Goals (SDGs), and identifying inequalities in health outcomes.

The credibility of WHO's data lies in its robust methodologies, strong partnerships, and commitment to transparency. The organization collaborates with governments, academic institutions, and international agencies to harmonize data collection processes, and it employs rigorous validation techniques to ensure accuracy. WHO updates its datasets regularly, making them a trusted resource for policymakers, researchers, and public health practitioners. Its annual reports, such as the "World Health Statistics," synthesize complex datasets into actionable insights, enabling stakeholders to address urgent health challenges and implement evidence-based solutions. By providing reliable, accessible, and actionable data, WHO empowers countries to make informed decisions that improve global health outcomes and reduce inequities.

#### World Bank :

The World Bank is a leading authority in the creation and provision of reliable, high-quality economic and development data, playing a critical role in global policymaking and research. As part of its mission to reduce poverty and promote sustainable development, the World Bank collects and analyzes data from over 200 countries, covering areas such as GDP, poverty rates, education, healthcare access, and infrastructure. This data is freely accessible through the World Bank Open Data platform, which ensures that stakeholders worldwide have access to the information needed to drive evidence-based decisions.

Among its most renowned resources is the World Development Indicators (WDI) database, which consolidates extensive datasets into a single, comprehensive tool for understanding global development trends. Other specialized datasets focus on issues like financial inclusion, gender equity, and climate change, offering granular insights that inform targeted interventions. The World Bank maintains strong partnerships with national statistical offices, academic institutions, and international organizations to ensure the reliability, timeliness, and relevance of its data. It employs standardized methodologies and advanced analytics to produce data that is consistent across countries and regions. By making these datasets publicly available, the World Bank not only enhances transparency and accountability but also fosters global collaboration to address pressing challenges, including poverty reduction, economic inequality, and climate resilience. Its commitment to providing robust, accessible data positions the institution as a vital resource for researchers, governments, and development agencies striving to build a better future.

## Presentation of the project
We have decided to test the environmental Kuznet's curve theory. The idea is that a country will pollute a lot for a while in order to develop, and then reach a sufficient level of economic development to decrease its carbon dioxyde emissions. We would then get an inverted u shaped line in the plan (GDP per capita, pollution). 

We wish to very the empirical accuracy of this theory, both by regions and periods of times. Does is apply to different economical structures ? Does it apply at any time over the years ? 

We will then consider the subject around a more human matter. Pollution to the cost of a population's health : an impact on GDP ? And if so, can we quantify it ? 

## Loading the datasets
```{r}
library(tidyr)
library(vroom)

DALYs <- vroom::vroom("DALYs.csv")
death_air_pollution <- vroom::vroom("death-air-pollution.csv")
SDG <- vroom::vroom("SDG indicators.csv")
GDP_per_capita <- vroom::vroom("GDPperCapita.csv", delim=",")
```
## Data cleaning

### GDP per capita
By trying to get the column names of this dataset, we observed that it did not load properly. There were the first four row of the Excel sheet presenting the data, hence compromising the understanding of the separator ",". We removed the first four rows that did not contain any data and got a proper dataset to work with. 
```{r}
GDP_per_capita_cleaned <- vroom::vroom("GDPperCapita.csv", skip = 4, delim=",") 
```
### DALYs
This dataset is hard to work with in the sense that the columns have non explicit names. hence the following changes. 

```{r}
library(dplyr)

DALYs <- DALYs %>%
  rename(
    Mean_Value = FactValueNumeric,
    lowest_value = FactValueNumericLow,
    highest_value = FactValueNumericHigh,
    Sex = Dim1,
    Country = Location
  )
```

### Death air pollution adn SDG
The observations are the same as for DALYs, hence the following changes.

```{r}
SDG <- SDG %>%
  rename(
    Mean_Value = FactValueNumeric,
    lowest_value = FactValueNumericLow,
    highest_value = FactValueNumericHigh,
    Sex = Dim1,
    Country = Location
  )
```

```{r}
death_air_pollution <- death_air_pollution %>%
  rename(
    Mean_Value = FactValueNumeric,
    lowest_value = FactValueNumericLow,
    highest_value = FactValueNumericHigh,
    Sex = Dim1,
    Country = Location
  )
```


