---
title: "Data Management Project"
author : Lou Dhiver and Lydie Lazure 
format: html
---
# Phase 2 of the project

### Github project :

https://github.com/Lydie03/datamana-DM14 

### Sources for data sets :

-   World Health Organization (WHO): Ambient air pollution attributable death rate (per 100,000 population, age-standardized) : https://www.who.int/data/gho/data/indicators/indicator-details/GHO/ambient-air-pollution-attributable-death-rate-(per-100-000-population-age-standardized)
-   World Health Organization (WHO): Ambient air pollution attributable DALYs (per 100,000 population, age-standardized) : https://www.who.int/data/gho/data/indicators/indicator-details/GHO/ambient-air-pollution-attributable-dalys-(per-100-000-population-age-standardized)
-   World Bank : GDP per capita (current US\$): https://data.worldbank.org/indicator/NY.GDP.PCAP.CD
-   World Health Organization (WHO): Ambient air pollution, Concentrations of fine particulate matter (PM2.5) per country : https://apps.who.int/gho/data/view.main.SDGPM25116v

#### World Health Organization (WHO) :

The World Health Organization (WHO) was created in 1948 and is responsible for public health on an international level. In particular, the organization provides health related data across the world. 

Through its Global Health Observatory (GHO), WHO offers an extensive repository of data on key health indicators, including life expectancy, maternal and child health, disease prevalence, vaccination coverage, and environmental factors like air and water quality. 

These will allow us to assess the accuracy of the Kuznets curve, in particular in terms of health and sustainability. 

#### World Bank :

The institution was created in 1945. It is an international financing institution that loans to developing countries, with the aim to provide economic development support. 

The World Bank provides a wide range of data for us to use. It covers over 200 countries with indicators such as GDP, education, gender inequality or infrastructure. 

## Presentation of the project
We have decided to test the environmental Kuznet's curve theory. The idea is that a country will pollute a lot for a while in order to develop, and then reach a sufficient level of economic development to decrease its carbon dioxyde emissions. We would then get an inverted u shaped line in the plan (GDP per capita, pollution). 

We wish to verify the empirical accuracy of this theory, both by region and time periods. Does it apply to different economic structures ? Does it apply at any time over the years ? 

We will then consider the subject from a more human perspective. Pollution and its cost to a population's health : does it impact GDP ? And if so, can we quantify it ? 

## Loading the datasets
```{r}
library(tidyr)
library(vroom)

DALYs <- vroom::vroom("DALYs.csv")
death_air_pollution <- vroom::vroom("death-air-pollution.csv")
SDG <- vroom::vroom("SDG indicators.csv")
GDP_per_capita <- vroom::vroom("GDPperCapita.csv", delim=",")
```
## Data cleaning

### GDP per capita
By trying to get the column names of this dataset, we observed that it did not load properly. There were the first four row of the Excel sheet presenting the data, hence compromising the understanding of the separator ",". We removed the first four rows that did not contain any data and got a proper dataset to work with. 
```{r}
GDP_per_capita_cleaned <- vroom::vroom("GDPperCapita.csv", skip = 4, delim=",") 
```
### DALYs
This dataset is hard to work with in the sense that the columns have non explicit names. hence the following changes. 

```{r}
library(dplyr)

DALYs <- DALYs %>%
  rename(
    Mean_Value = FactValueNumeric,
    lowest_value = FactValueNumericLow,
    highest_value = FactValueNumericHigh,
    Sex = Dim1,
    Country = Location
  )
```

### Death air pollution and SDG
The observations are the same as for DALYs, hence the following changes.

```{r}
SDG <- SDG %>%
  rename(
    Mean_Value = FactValueNumeric,
    lowest_value = FactValueNumericLow,
    highest_value = FactValueNumericHigh,
    Sex = Dim1,
    Country = Location
  )
```

```{r}
death_air_pollution <- death_air_pollution %>%
  rename(
    Mean_Value = FactValueNumeric,
    lowest_value = FactValueNumericLow,
    highest_value = FactValueNumericHigh,
    Sex = Dim1,
    Country = Location
  )
```


## Data description
To answer our research question, we will need data around pollution, GDP per capita and health. 
To this end, we will be working with databases from the World Health Organization, and the World Bank. 

References are listed in the appendix. 

### DALYs
This aggregated indicator is extracted from the World Health Organization database. It is described as follows : "DALYs for a disease or health condition are the sum of the years of life lost to due to premature mortality (YLLs) and the years lived with a disability (YLDs) due to prevalent cases of the disease or health condition in a population". 
We will mostly use the variable cross countries, classified into regions (Africa, Eastern Europe, Asia, North and South America), by sex (male/female), by year (2010 to 2019) and by cause of death (cancer, trachea etc.). 
This will allow us to evaluate possible external factors, such as armed conflicts, that might have had an impact on specific regions during a specific year. 

This dataset will then be compared to pollution to answer our final sub research question : can a country observe long term growth while polluting despite the negative externalities in terms of health ? 

```{r}
# Data summary
c("Number Of Countries", "Mean DALY", "Median DALY", "Standard deviation", "DALY variable type")
c(lenght(DALYs$Country), mean(DALYs$Indicator), median(DALYs$Indicator, na.rm=FALSE), sd(DALYs$Indicator, na.rm=FALSE), typeof(DALYs$Indicator))

table(DALYs$Country, DALYs$Period, DALYs$Mean_Value) ## Still searching for better answer

# Data graphical representations

library(ggplot2)

# Grouping mean over time per country
DALYs_unique <- DALYs %>%
  group_by(ParentLocation) %>%
  summarise(Mean_Value = mean(Mean_Value, na.rm = TRUE))

# Graphical representation
ggplot(data = DALYs_unique, aes(x = ParentLocation, y = Mean_Value, fill = ParentLocation)) +
  geom_bar(stat = "identity", color = "black") + 
  labs(
    title = "Distribution of Mean DALYs by Parent Location",
    x = "Parent Location",
    y = "Mean DALYs (Disability-Adjusted Life Years)"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_brewer(palette = "Set3") 
```
When we group the mean of DALYs values over time per country, we get a rather interesting picture of the situation cross regions. It seems DALYs are prevalent in Africa, Eastern Mediterranean and South-East Asia. This will allow us to define control groups and to consider levels of development compared to DALYs. 

### Death air pollution
This data set will be used to complete DALYs. It also comes from the World Health Organization and is organized with the same categories and classifications. 
It will allow us to get a more precise analysis on the effect of carbone emissions on health. 

```{r}
head(death_air_pollution)
summary(death_air_pollution)

library(ggplot2)

# Grouping mean over time per country
death_air_pollution_unique <- death_air_pollution %>%
  group_by(ParentLocation) %>%
  summarise(Mean_Value = mean(Mean_Value, na.rm = TRUE))

# Graphical representation
ggplot(data = death_air_pollution_unique, aes(x = ParentLocation, y = Mean_Value, fill = ParentLocation)) +
  geom_bar(stat = "identity", color = "black") + 
  labs(
    title = "Distribution of mean deaths due to air pollution by Parent Location",
    x = "Parent Location",
    y = "Mean deaths due to air pollution"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_brewer(palette = "Set3") 
```
Here it seems deaths due to air pollution is prevalent in South East Asia and Western Pacific. 

### SDG
This dataset includes the measurement of fine particulate matter, that is to say the concentration of toxic elements in the air. Again, since the dataset comes from the World Health organization, the classification remains the same. It will allow us to get a cross country analysis, from 2010 to 2019.
We will compare the degree to which some regions are polluted and their GDP over the year.

```{r}
head(SDG)
summary(SDG)

library(ggplot2)

# Grouping mean over time per country
SDG_unique <- SDG %>%
  group_by(ParentLocation) %>%
  summarise(Mean_Value = mean(Mean_Value, na.rm = TRUE))

# Graphical representation
ggplot(data = SDG_unique, aes(x = ParentLocation, y = Mean_Value, fill = ParentLocation)) +
  geom_bar(stat = "identity", color = "black") + # Barres remplies
  labs(
    title = "Average Fine Particulate Matter by Country (2010-2019)",
    x = "Parent Location",
    y = "Average Fine Particulate Matter (µg/m³)"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_fill_brewer(palette = "Set3")
```
The pattern here is quite similar to DALYs. SDG is prevalent in Africa, Eastern Mediterranean and South-East Asia. 

### GDP per capita
GDP (Gross domestic product) is obtained with the following macroeconomics formula :
`GDP=C+I+G+(X−M)`
That is to say the aggregation of consumption (C), investments (I), government spending (G) and the commercial balance (exports minus imports).
GDP per capita is the previous aggregate divided by the population size. This choice will allow us to correct for countries size bias. 

GDP will be our wealth indicator in the simulation of the Kuznet's environmental curve. 

```{r}
head(GDP_per_capita_cleaned)
summary(GDP_per_capita_cleaned)

library(ggplot2)

## Filtering by region
rows_to_keep <- c(182, 216, 231, 232, 237, 241, 260)
GDP_per_capita_cleaned_filtered <- GDP_per_capita_cleaned |> 
  slice(rows_to_keep)

## Calculating mean over the specified period for each region
mean_values <- rowMeans(GDP_per_capita_cleaned_filtered[, 5:50], na.rm = TRUE) 
GDP_per_capita_cleaned_filtered$Mean_5_to_50 <- mean_values 

## Creating a bar chart with ggplot2
library(ggplot2)
ggplot(data = GDP_per_capita_cleaned_filtered, aes(x = `Country Name`, y = Mean_5_to_50)) +
  geom_bar(stat = "identity", fill = "steelblue") + 
  labs(
    x = "Country Name", 
    y = "Average GDP per Capita (Columns 5-50)", 
    title = "Average GDP per Capita by Country"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1) 
  )
```
We chose to use the aggregations made by the world bank to get a broad view of GDP per capita across the world. We can draw in that sense our first intuitions. It seems regions of the world have a very unequal distribution of wealth, especially compared to the OECD countries. Low incomes seem to match low revenue activities such as agriculture or textile industries, when the OECD countries match group services and technologies. Hence supposedly a higher pollution rate induced by low income countries prevalent economical activities.If we refer to our previous graphical representations, we can note however that it is not the case for Africa : agriculture does not pollute as much as textile industries in The East and Pacific Asia. 

*ZIP on Moodle*

